Hydras \& Co.

### 2022-09-06 doc/chapter-encoding.tex
start chapter on goedel encodings

>>>>

      \subsubsection{Main properties}

      In order to prove that the function \texttt{cPair} is primitive recursive, we express it as a composition of already proven primitive recursive functions.

      \inputsnippets{cPair/cPairIsPR}

      \texttt{cPair}'s injectivity is stated by two lemmas:

      \inputsnippets{cPair/cPairInj1, cPair/cPairInj2}

      \subsection{Projections}

      Let $a$ be some natural number, we look for two natural numbers $x$ and $y$ such that $\texttt{cPair}\,x\,y=a$. More we want to prove that the function which compute $x$ [resp. $y$] out of $a$ are primitive recursive.

      Let us show on a small example how these projections are defined. Let's take for instance $a=11$. Please look at the diagram of Fig.~\ref{fig:cpair}.

      The following function looks for the anti-diagonal $\{(x,y)|x+y=k\}$ the number $11$ belongs to, and returns $k$.

      \inputsnippets{cPair/searchXY}

      In our example, $11$ belongs to the anti-diagonal $\{(x,y)|x+y=4\}$, which contains values from $10$ to $14$.
      Thus the line containing $11$ is the line $x=11-10=1$, and the column is $y=4-1= 3$. Finally, we get  $11=\texttt{cPair}\,1\,3$.

       \inputsnippets{cPair/cPairPI12}

       The following lemmas (still from \href{../theories/html/hydras.Ackermann.cPair.html}{Ackermann.cPair}) prove the correctness of these projections.

       \inputsnippets{cPair/cPairProjections,
         cPair/cPairProjections1, cPair/cPairProjections2,
         cPair/cPairPi1IsPR, cPair/cPairPi2IsPR}

       Finally, let us show a few inequalities.

       \inputsnippets{cPair/cPairLe1, cPair/cPairLe1A,
         cPair/cPairLe2, cPair/cPairLe2A, cPair/cPairLe3,
         cPair/cPairLt1,cPair/cPairLt2}
